---
title: "Untitled"
author: "Kiera Matiska"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(tmap)
library(sf)
library(janitor)
library(plotly)
library(beepr)
```


```{r}
county_sf <- read_sf(here("US_County_Boundaries copy/US_County_Boundaries.shp")) %>% 
  clean_names()

county_subset_sf <- county_sf %>% 
  select(state, county, shape_area)
```

```{r}
county_subset_sf %>% st_crs()
```

```{r}
x <- ggplot(data = county_subset_sf) +
  geom_sf(aes(fill = state), color = "white", size = 0.1)

# tmap(county_subset_sf)

# beepr::beep(6)
```

```{r}
co_county <- county_subset_sf %>% 
  filter(state == "Alabama")

county_state_sf <- county_subset_sf %>% 
  filter(state == "Alaska")

# ggplot(data = county_state_sf) +
#   geom_sf(aes(fill = county, color = "white", size = 0.1))
# 
# tmap_mode(mode = "view")
# 
# tm_shape(county_state_sf) +
#   tm_fill("county") +
#   tm_borders(col = "black")
# 
# beepr::beep(6)
```



```{r}
food_access <- read_csv("food_access_subset copy.csv") %>% 
  drop_na() %>% 
  mutate(income = median_family_income)
```

income_snap_table <- reactive({
    food_access %>% 
      filter(income == input$median_family_income, state == input$state) %>% 
      mutate(median_family_income = case_when(
        median_family_income >= "0" ~ "1",
        median_family_income >= "25000" ~ "2",
        median_family_income >= "50000" ~ "3",
        median_family_income >= "75000" ~ "4",
        median_family_income >= "100000" ~ "5"
      ))
      
      
```{r}
income_snap_table <- food_access %>% 
  mutate(income = case_when(
    income >= "0" & income < "25000" ~ "1",
    income >= "25000" & income < "50000" ~ "2",
    income >= "50000" & income < "75000" ~ "3",
    income >= "75000" & income < "100000" ~ "4",
    income >= "100000" & income < "125000" ~ "5",
    income >= "125000" & income < "150000" ~ "6",
    income >= "150000" & income < "175000" ~ "7",
    income >= "175000" & income < "200000" ~ "8",
    income >= "200000" & income < "225000" ~ "9",
    income >= "225000" & income <= "250000" ~ "10"
  )) %>% 
  group_by(state, income) %>% 
  summarize(total_pop = sum(pop2010),
            sum_snap = sum(tract_snap))
```

Bargraph displaying which the population count for different ethnicities that are beyond 10 miles from a supermarket

Ethnicities:

- White 
- African American
- Asian
- Native Hawaiian or other Pacific Islander
- American Indian or Alaska Native
- Hispanic
- Other

```{r}
# white
eth_white_sub <- food_access %>% 
  select(state, urban, lawhitehalf, lawhite1, lawhite10, lawhite20) %>% 
  group_by(state) %>% 
  summarize(sum_white_half = sum(lawhitehalf),
            sum_white_1 = sum(lawhite1),
            sum_white_10 = sum(lawhite10),
            sum_white_20 = sum(lawhite20))

# black
eth_black_sub <- food_access %>% 
  select(state, urban, lablackhalf, lablack1, lablack10, lablack20) %>% 
  group_by(state) %>% 
  summarize(sum_black_half = sum(lablackhalf),
            sum_black_1 = sum(lablack1),
            sum_black_10 = sum(lablack10),
            sum_black_20 = sum(lablack20))

#asian
eth_asian_sub <- food_access %>% 
  select(state, urban, laasianhalf, laasian1, laasian10, laasian20) %>% 
  group_by(state) %>% 
  summarize(sum_asian_half = sum(laasianhalf),
            sum_asian_1 = sum(laasian1),
            sum_asian_10 = sum(laasian10),
            sum_asian_20 = sum(laasian20))

# Native Hawaiian or other Pacific Islander
eth_nhopi_sub <- food_access %>% 
  select(state, urban, lanhopihalf, lanhopi1, lanhopi10, lanhopi20) %>% 
  group_by(state) %>% 
  summarize(sum_nhopi_half = sum(lanhopihalf),
            sum_nhopi_1 = sum(lanhopi1),
            sum_nhopi_10 = sum(lanhopi10),
            sum_nhopi_20 = sum(lanhopi20))

# American Indian or Alaska Native
eth_aian_sub <- food_access %>% 
  select(state, urban, laaianhalf, laaian1, laaian10, laaian20) %>% 
  group_by(state) %>% 
  summarize(sum_aian_half = sum(laaianhalf),
            sum_aian_1 = sum(laaian1),
            sum_aian_10 = sum(laaian10),
            sum_aian_20 = sum(laaian20))

# Hispanic
eth_hisp_sub <- food_access %>% 
  select(state, urban, lahisphalf, lahisp1, lahisp10, lahisp20) %>% 
  group_by(state) %>% 
  summarize(sum_hisp_half = sum(lahisphalf),
            sum_hisp_1 = sum(lahisp1),
            sum_hisp_10 = sum(lahisp10),
            sum_hisp_20 = sum(lahisp20))

eth_omultir_sub <- food_access %>% 
  select(state, urban, laomultirhalf, laomultir1, laomultir10, laomultir20) %>% 
  group_by(state) %>% 
  summarize(sum_omultir_half = sum(laomultirhalf),
            sum_omultir_1 = sum(laomultir1),
            sum_omultir_10 = sum(laomultir10),
            sum_omultir_20 = sum(laomultir20))
```

merge the two data sets

```{r}
# determine whether each county is more urban or rural
rur_urb_al <- 

rur_urb_food <- food_access %>% 
  group_by(state, county) %>% 
  summarize(mean_urban = mean(urban),
            total_pop = sum(pop2010))

rur_urb_type <- rur_urb_food %>% 
  mutate(type = case_when(
    mean_urban < "0.5" ~ "Rural",
    mean_urban >= "0.5" ~ "Urban"
  ))
```


```{r}
county_subset_df <- county_subset_sf %>% 
  as.data.frame()

sf_merge <- merge(rur_urb_type, county_subset_df)

rur_urb_geom_sf <- st_sf(sf_merge)

# ca_counties_df <- ca_counties_sf %>% 
#   as.data.frame() %>% 
#   select(-geometry)
# only use this if you need to make a data frame
```

Export to shape file

```{r}
st_write(nc, paste0(tempdir(), "/", "nc.shp"))

st_write(rur_urb_geom_sf, "rur_urb_geom_sf.shp")

st_write(rur_urb_geom_sf, file = "rur_urb_geom_sf.shp")
```

## New subset for widget 4

These columns:

- LATracts_half
- LATracts1
- LATracts10
- LATracts20
- LATractsVehicle_20
- lapophalf
- lahunvhalf
- lapop1
- lahunv1
- lapop10
- lahunv10
- lapop20
- lahunv20
- TractHUNV

```{r}
vehicle_food <- food_access %>% 
  select(state, county, urban, pop2010, lapophalf, lapop1, lapop10, lapop20, starts_with("lahunv"), tract_hunv)

vehicle_food_urb <- vehicle_food %>% 
  mutate(urban = case_when(
    urban == "1" ~ "Urban",
    urban == "0"~ "Rural"
  ))

vehicle_food_sum <- vehicle_food_urb %>% 
  group_by(state, county, urban) %>% 
  summarize(tot_pop = round(sum(pop2010), 0),
            pop_half = round(sum(lapophalf), 0),
            pop1 = round(sum(lapop1), 0),
            pop10 = round(sum(lapop10), 0),
            pop20 = round(sum(lapop20), 0),
            vehicle_half = round(sum(lahunvhalf), 0),
            vehicle1 = round(sum(lahunv1), 0),
            vehicle10 = round(sum(lahunv10), 0),
            vehicle20 = round(sum(lahunv20), 0),
            sum_tract = sum(tract_hunv))

write_csv(vehicle_food_sum, "vehicle_food_subset.csv")
```

```{r}
pivot_longer_vehicle <- vehicle_food_sum %>% 
  pivot_longer(vehicle_half:vehicle20) %>% 
  select(-sum_tract)
```

# New Ethnicity Subset

- White 
- African American
- Asian
- Native Hawaiian or other Pacific Islander
- American Indian or Alaska Native
- Hispanic
- Other






      
      
      
      

